---
title: "Weekly Guide 3"
linktitle: "Week 3"
citeproc: true
bibliography: ../../static/bib/BibTeX_Library.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
menu:
  content:
    parent: Weekly Guides
    weight: 3
type: docs
output:
  blogdown::html_page:
    toc: false
---

```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(error = TRUE, comment = NA, warnings = FALSE, errors = FALSE, messages = FALSE, tidy = FALSE, echo = FALSE)
```

This week we'll talk about risk premiums and examine demand for health insurance (i.e., willingness to pay for health insurance).

### Objectives for this week
- Explain and draw graphically how the risk premium varies with the curvature of the utility function, probability of illness, and amount of potential loss
- Calculate risk premium and willingness-to-pay for health insurance
- Discuss viability of high-risk pools
 
### Activities
- In-class problem on the demand for health insurance. The worksheet is [here](/resource/wksheet2.html).
- Second [weekly check-in](/assignment/weekly-check-in.html) due on Friday.

### Supplemental readings
- Chapter 4 of @sloan2012

### Pre-class supplemental materials

**Day 3: Monday, January 24, 2022**


<center>

```{r}
knitr::include_url("https://imccart.github.io/econ372-content/01-Insurance/preclass-03.html")
```

<p><a class="btn btn-primary btn-lg" target="_blank" href="https://imccart.github.io/econ372-content/01-Insurance/preclass-03.html"><i class="fas fa-external-link-alt"></i> View all slides in new window</a>

</center>
