---
title: "Weekly Guide 2"
linktitle: "Week 2"
citeproc: true
bibliography: ../../static/bib/BibTeX_Library.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
menu:
  content:
    parent: Weekly Guides
    weight: 2
type: docs
output:
  blogdown::html_page:
    toc: false
---

```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(error = TRUE, comment = NA, warnings = FALSE, errors = FALSE, messages = FALSE, tidy = FALSE, echo = FALSE)
```

Today is all about risk and uncertainty. We will eventually use these concepts to study insurance markets and premiums.

### Objectives
- Calculate expected values and expected utility
- Define a risk pool
- Define risk aversion and diminishing marginal utility
 
### Activities
- Discussion of Humana and importance of risk pools
- In-class example on expected values
- In-class problem on expected costs, expected utilities, and risk aversion. The worksheet is [here](/resource/wksheet1.html).
- First [weekly check-in](/assignment/weekly-check-in.html) due on Friday.

### Supplemental readings
- Sections 5.1-5.2 of @pindyck2018

### Pre-class supplemental materials

**Day 2: Wednesday, January 19, 2022**


<center>

```{r}
knitr::include_url("https://imccart.github.io/econ372-content/01-Insurance/preclass-02.html")
```

<p><a class="btn btn-primary btn-lg" target="_blank" href="https://imccart.github.io/econ372-content/01-Insurance/preclass-02.html"><i class="fas fa-external-link-alt"></i> View all slides in new window</a>

</center>
